Spatial variation of the native colon microbiota in healthy adults

### 2018-01-30 submission/manuscript.Rmd
discussion updated revision

The distribution of these disease-associated microbes in healthy patients is of interest their presence can be predictive of disease prior to the onset of symptoms @Baxter2016. Spatial organization of _Fusobacteria_ and other bacterial species into polymicrobial biofilms that can invade the gut mucosa have been linked CRC (cite new paper). The organization into biofilms promotes tumorigenesis by allowing bacteria to grow near the epithelium, promoting inflammation, genotoxicity and metabolic changes that favor tumor cell growth (cite new paper and kaitlin). In one study of CRC biofilms and tumors, all of the proximal tumors examined contained a polymicrobial biofilm on the tumor mucosa @Dejea2014. Further examination of these tumors identified _Fusobacteria_ species as members of the biofilm. These results indicate that it is not only the presence of the bacterium but the tumor community as a whole that contributes to tumorigenesis @Dejea2014. Driver passeneger model (cite new paper and KF).  Need to study healthy people/biofilms more to understand how can affect disease. Further analysis of the distribution and activities of these pathogens along the colon is needed to elucidate a mechanism for development of CRC or IBD subtypes in the proximal or distal colon.

paragraph on bacterial organization/spatial variation along colon vs at tumor site/biofilms

>>>>

Spatial organization of _Fusobacteria_ and other bacterial species into polymicrobial biofilms that can invade the gut mucosa have been linked CRC (cite new paper). The organization into biofilms promotes tumorigenesis by allowing bacteria to grow near the epithelium, promoting inflammation, genotoxicity and metabolic changes that favor tumor cell growth (cite new paper and kaitlin). In one study of CRC biofilms and tumors, all of the proximal tumors examined contained a polymicrobial biofilm on the tumor mucosa @Dejea2014. Further examination of these tumors identified _Fusobacteria_ species as members of the biofilm. These results indicate that it is not only the presence of the bacterium but the tumor community as a whole that contributes to tumorigenesis @Dejea2014. A 'driver-passenger' model has been proposed as a mechanism for biofilm assembly in the gut (Tjalsma 2012, flynn). In this model, 'driver' species such as _Fusobacterium spp_ and _Porphyromonas spp_ exert tumorigenic effects locally and create a niche for colonization or adherence of 'passenger' species that comprise the rest of the biofilm (Tjalsma and Flynn). Thus the the distribution of these disease-associated microbes in healthy patients is of interest as their presence can be predictive of disease prior to the onset of symptoms @Baxter2016. A better understanding of the early microbial changes in the gut microboiome is essential for elucidating a mechanism for development of CRC or IBD subtypes in the proximal or distal colon.
